---
title: Checking annotation status
subtitle: Bird song evolution
author: <a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a>
date: "`r Sys.Date()`"
toc: true
toc-depth: 2
toc-location: left
number-sections: true
highlight-style: pygments
format:
  html:
    df-print: kable
    code-fold: true
    code-tools: true
    code-copy: true
    embed-resources: true
editor_options: 
  chunk_output_type: console
---


<!-- this code add line numbers to code blocks -->
<!-- only works when code folding is not used in yaml (code_folding: show) -->

```{=html}
<style>
body
  { counter-reset: source-line 0; }
pre.numberSource code
  { counter-reset: none; }
</style>
```

```{r set root directory, echo = FALSE}

# set working directory as project directory or one directory above,
rootdir <- try(rprojroot::find_rstudio_root_file(), silent = TRUE)
if (is(rootdir, "try-error")) rootdir <-  ".."
knitr::opts_knit$set(root.dir = rootdir)

```

```{r add link to github repo, echo = FALSE, results='asis'}

# print link to github repo if any
if (file.exists("./.git/config")){
  config <- readLines("./.git/config")
  url <- grep("url",  config, value = TRUE)
  url <- gsub("\\turl = |.git$", "", url)
  cat("\nSource code, data and annotation protocol found at [", url, "](", url, ")", sep = "")
  }

```

```{r setup style, echo = FALSE, message = FALSE, warning=FALSE}

# options to customize chunk outputs
knitr::opts_chunk$set(
  class.source = "numberLines lineAnchors", # for code line numbers
  tidy.opts = list(width.cutoff = 65), 
  tidy = TRUE,
  message = FALSE,
  warning = FALSE
 )

```

<!-- skyblue box -->
<div class="alert alert-info">

# Purpose {.unnumbered .unlisted}

- Double-check annotations

</div>

&nbsp; 



# Load packages {.unnumbered .unlisted}

```{r load packages}

# knitr is require for creating html/pdf/word reports
# formatR is used for soft-wrapping code

# install/ load packages
sketchy::load_packages(packages = c("knitr", "formatR", "rprojroot", "viridis", "googlesheets4", "warbleR", "maRce10/Rraven", "googledrive", "maRce10/ohun"))

source("~/Dropbox/R_package_testing/Rraven/R/imp_raven.R")

source("~/Dropbox/R_package_testing/Rraven/R/internal_functions.R")
```

# Consolidate sound files
```{r, eval = FALSE}

cns <- consolidate(path = "/media/m/Campylopterus/bird_song_evolution/by_family", dest.path = "/run/user/1000/gvfs/smb-share:server=cinnas.local,share=neurobiología/marcelo_araya/bird_song_evolution/consolidated_sound_files", parallel = 10)

```


# Update annotations from google drive

```{r update annotations, eval = FALSE}

# read data
ann_folder <- drive_find(pattern = "raven_selections", type = "folder")

# list family folders
fam_folders <- drive_ls(ann_folder)

# list files
anns_ls <- lapply(1:nrow(fam_folders), function(x)
    drive_ls(fam_folders$id[x])
    )

all_ans <- do.call(rbind, anns_ls)

all_ans$modified.time <- sapply(1:nrow(all_ans), function(x) 
all_ans$drive_resource[[x]]$modifiedTime)

all_ans <- all_ans[, c("name", "id", "modified.time")]


local_files <- list.files("./data/processed/annotations", pattern = ".txt", full.names = FALSE)
local_file_max_time <- max(file.mtime(list.files("./data/processed/annotations", pattern = ".txt", full.names = TRUE)))

# keep those changed after older local file
all_ans <- all_ans[all_ans$modified.time > local_file_max_time | !all_ans$name %in% local_files,]

owd <- getwd()

setwd("./data/processed/annotations")

# copy locally new files found in google drive 
new_files <- sapply(1:nrow(all_ans), function(x) {
    
    drive_download(all_ans$id[x], overwrite = TRUE)
     
    })

xz vcsetwd(owd)  

new_file_count <- nrow(all_ans)

if (new_file_count > 0)
    print(paste(new_file_count, "new annotations file(s) found"))

```


# Descriptive stats
```{r descriptive stats, eval = TRUE}

rec_data <- read_sheet('https://docs.google.com/spreadsheets/d/16ukhyf37hm13f1FXB2JQ-tCfHoxXy2qj_TX2xWvTVu8/edit#gid=148139271')

rec_data$assigned_to[rec_data$assigned_to == "NA"] <- NA
anns <- imp_raven(path = "./data/processed/annotations",  all.data = TRUE, warbler.format = TRUE, name.from.file = TRUE, ext.case = "lower", pb = FALSE)

ann_files <- unique(c(anns$selec.file, unlist(.Options$Rraven)))

ann_files <- data.frame(file = ann_files)

ann_files$species <- gsub("_", " ", sapply(strsplit(ann_files$file, "-"), "[", 1))


# all(ann_files$species %in% rec_data$species)

ann_files$family <- sapply(ann_files$species,  function(x) rec_data$family[rec_data$species == x][1])

# get those in which the name doesnt match the expected
weird_files <- ann_files[!grepl("Table.1.selections.txt", ann_files$file) | ann_files$file %in% unlist(.Options$Rraven), ]

weird_files$problem <- "file name"
weird_files$problem[weird_files$file %in% unlist(.Options$Rraven)] <- "Empty file"

# keep only those with "good" names
ann_files <- ann_files[!ann_files$file %in% weird_files$file, ]

# counts per family
fam_count <- aggregate(species ~ family, data = rec_data, length)
names(fam_count)[2] <- "total"

fam_count$annotated <- sapply(fam_count$family, function(x)
    sum(ann_files$family == x))

fam_count$prop.annotated  <- round(fam_count$annotated / fam_count$total, 2)

fam_count$assinged.to <- sapply(fam_count$family, function(x)
    paste(unique(rec_data$assigned_to[rec_data$family == x]), collapse = "/"))

fam_count$assinged.to[fam_count$assinged.to == "NA"] <- "not assigned"

weird_files$assinged.to <-  sapply(weird_files$family, function(x)
    paste(unique(rec_data$assigned_to[rec_data$family == x]), collapse = "/"))

```

- `r nrow(ann_files)` species from `r length(unique(ann_files$family))` families already annotated (`r  paste0(round(nrow(ann_files) / nrow(rec_data), 2) * 100, "%")` of all recordings)

- `r sum(fam_count$prop.annotated >= 0.9)` families with 90% of the recordings annotated (`r round(sum(fam_count$prop.annotated >= 0.9) / nrow(fam_count), 2) * 100` of all families)

- `r sum(fam_count$prop.annotated <= 0.05)` families with less than 5% of the recordings annotated (`r round(sum(fam_count$prop.annotated <= 0.05) / nrow(fam_count), 2) * 100` of all families)

- `r sum(weird_files$problem == "file name")` weirdly named file(s) and `r sum(weird_files$problem == "Empty file")` empty file(s):
```{r}

weird_files

```


# Species per family

- `prop.annotated`: proportion of species already annotated (green > 0.9, yellow > 0.7, red  < 0.5)
- Green rows highlight families in which all available species were already annotated

```{r, eval = TRUE}

prop_analzyed <- fam_count$prop.annotated
fam_count$prop.annotated <-
      ifelse(
       fam_count$prop.annotated > 0.9,
        kableExtra::cell_spec(
          fam_count$prop.annotated,
          "html",
          color = "white",
          background = "green",
          bold = TRUE,
          font_size = 12,
          new_tab = TRUE
        ),
        ifelse(
       fam_count$prop.annotated > 0.5,
        kableExtra::cell_spec(
          fam_count$prop.annotated,
          "html",
          color = "black",
          background = "yellow",
          bold = TRUE,
          font_size = 12,
          new_tab = TRUE
        ),
        kableExtra::cell_spec(
          fam_count$prop.annotated,
          "html",
          color = "white",
          background = "red",
          bold = TRUE,
          font_size = 12,
          new_tab = TRUE
        )
       )
      )

fam_count$assinged.to <-
      ifelse(
       fam_count$assinged.to == "not assigned",
        kableExtra::cell_spec(
          fam_count$assinged.to,
          "html",
          color = "white",
          background = "red",
          bold = TRUE,
          font_size = 12,
          new_tab = TRUE
        ),
         kableExtra::cell_spec(fam_count$assinged.to, "html", new_tab = TRUE)
       )


fam_count$recordings <- 
  kableExtra::cell_spec("link", "html", link = paste0("https://6f33fa7f78ea46e2aaca-my.sharepoint.com/personal/marcelo_araya_ucr_ac_cr/_layouts/15/onedrive.aspx?ga=1&id=%2Fpersonal%2Fmarcelo%5Faraya%5Fucr%5Fac%5Fcr%2FDocuments%2Fbird%5Fsong%5Frecordings%2F", fam_count$family), new_tab = TRUE)

fam_count$spectrograms <- ifelse(prop_analzyed > 0,
  kableExtra::cell_spec("link", "html", link = paste0("https://6f33fa7f78ea46e2aaca-my.sharepoint.com/personal/marcelo_araya_ucr_ac_cr/_layouts/15/onedrive.aspx?ga=1&id=%2Fpersonal%2Fmarcelo_araya_ucr_ac_cr%2FDocuments%2Fannotated_spectrograms%2F", fam_count$family), new_tab = TRUE), "")


fam_count_kbl <- kableExtra::kbl(
      fam_count,
      row.names = FALSE,
      escape = FALSE,
      format = "html",
      digits =
    )

  fam_count_kbl <-
    kableExtra::row_spec(
      kable_input = fam_count_kbl,
      row =  which(prop_analzyed == 1),
      background = grDevices::adjustcolor("#6DCD59FF", alpha.f = 0.3)
    )

  fam_count_kbl <-
    kableExtra::kable_styling(
      fam_count_kbl,
      bootstrap_options = c("striped", "hover", "condensed", "responsive"),
      full_width = FALSE,
      font_size = 12
    )

 fam_count_kbl

```


# Check annotations
```{r, eval = TRUE}

warbleR_options(wav.path = "/run/user/1000/gvfs/smb-share:server=cinnas.local,share=neurobiología/marcelo_araya/bird_song_evolution/consolidated_sound_files")

# check_sound_files()

anns <- anns[, c("sound.files", "selec", "start", "end", "bottom.freq", "top.freq", "selec.file", "element", "song")]

anns <- anns[anns$selec.file %in% ann_files$file, ]

anns$species <- gsub("_", " ", sapply(strsplit(anns$selec.file, "-"), "[", 1))

anns$family <- sapply(anns$species,  function(x) rec_data$family[rec_data$species == x][1])

anns$assinged.to <-  sapply(anns$family, function(x)
    paste(unique(rec_data$assigned_to[rec_data$family == x]), collapse = "/"))

# all(ann_files$species %in% rec_data$species)

cs <- check_sels(anns, parallel = 20, pb= FALSE)

# table(cs$check.res[cs$check.res != "OK"])

# unique(cs$sound.files[cs$check.res == "sound file not found"])

# exp_raven(cs, path = "./data/processed", sound.file.path =  "/media/m/Seagate Portable Drive/bird_song_recordings/consolidated_files",file.name = "combined_annotations_22-03-2023")

ok_anns <- cs[cs$check.res == "OK", ]

# feature_acoustic_data(.Options$warbleR$path)

feature_bysong <- feature_reference(reference = ok_anns, by.sound.file = TRUE)

```

- Check the following annotations:

```{r, eval = TRUE}

problematic_sels <- cs[cs$check.res != "OK", ]

# weird duration or frequency range
long_anns <- cs[cs$end - cs$start > 3 & !cs$family %in% c("Tinamidae", "Accipitridae", "Falconidae"), ] 
long_anns$check.res <- "longer than 3 s"

broad_anns <- cs[cs$top.freq - cs$bottom.freq > 10, ] 
broad_anns$check.res <- "frequency range wider than 10 kHz"

song_na <- cs[is.na(cs$song), ] 
song_na$check.res <- "NAs in 'song'"

elem_na <- cs[is.na(cs$element), ] 
elem_na$check.res <- "NAs in 'element'"

problematic_sels <- rbind(problematic_sels, long_anns, broad_anns, song_na, elem_na)

problematic_sels <- problematic_sels[, c("sound.files","family", "assinged.to",  "selec", "check.res")]

problematic_sels$spectrograms <- ""
names(problematic_sels) <- c("sound.file", "family", "assinged.to", "selection", "problem", "spectrograms")

problematic_sels <- problematic_sels[order(problematic_sels$family, problematic_sels$sound.file, problematic_sels$selection), ]

sf <- gsub("\\.wav", "", problematic_sels$sound.file)
sf <- gsub("_", "%5F", sf)
sf <- gsub("-", "%2D", sf)

problematic_sels <- problematic_sels[order(problematic_sels$sound.file, problematic_sels$selection), ]

problematic_sels$spectrograms <- ifelse(file.exists(file.path("./data/processed/annotated_spectrograms", problematic_sels$family, gsub(".wav", "--p1.jpeg",problematic_sels$sound.file))),  
  kableExtra::cell_spec("link", "html", link = paste0("https://6f33fa7f78ea46e2aaca-my.sharepoint.com/personal/marcelo_araya_ucr_ac_cr/_layouts/15/onedrive.aspx?q=",gsub("_", "%5F", sapply(strsplit(problematic_sels$sound.file, "-"), "[", 1)), "&view=7&id=%2Fpersonal%2Fmarcelo%5Faraya%5Fucr%5Fac%5Fcr%2FDocuments%2Fannotated%5Fspectrograms&searchScope=folder"), new_tab = TRUE), "")

problematic_sels$recording <-    
   kableExtra::cell_spec("link", "html", link = paste0("https://6f33fa7f78ea46e2aaca-my.sharepoint.com/personal/marcelo_araya_ucr_ac_cr/_layouts/15/onedrive.aspx?q=",gsub("_", "%5F", sapply(strsplit(problematic_sels$sound.file, "-"), "[", 1)), "&view=7&id=%2Fpersonal%2Fmarcelo%5Faraya%5Fucr%5Fac%5Fcr%2FDocuments%2Fbird%5Fsong%5Frecordings&searchScope=folder"), new_tab = TRUE)
    

problematic_sels_kbl <- kableExtra::kbl(
      problematic_sels,
      row.names = FALSE,
      escape = FALSE,
      format = "html",
      digits = 3
)

  problematic_sels_kbl <-
    kableExtra::kable_styling(
      problematic_sels_kbl,
      bootstrap_options = c("striped", "hover", "condensed", "responsive"),
      full_width = FALSE,
      font_size = 12
    )

 problematic_sels_kbl

```

## Created annotated spectrograms
```{r, eval = FALSE}

ann_spec_path <- "./data/processed/annotated_spectrograms"

# create folders for each family
for(i in unique(rec_data$family))
    if (!file.exists(file.path(ann_spec_path, i)))
    dir.create(file.path(ann_spec_path, i))

out <- warbleR:::pblapply_wrblr_int(sort(unique(cs$sound.files)), pbar = TRUE, cl = 8, function(x)
    {
    sub_anns <- cs[cs$sound.files == x, ]
    frq_range <- range(c(sub_anns$bottom.freq, sub_anns$top.freq))
    
    frq_range[1] <- frq_range[1] - 1
    if (frq_range[1] < 0) frq_range[1] <- 0

    frq_range[2] <- frq_range[2] + ((frq_range[2] - frq_range[1]) / 3)
    if(frq_range[2] < 6) frq_range[2] <- 6
    
    if (frq_range[2] > 22.05) frq_range[2] <- 22.05
    
    sub_anns$selec <- paste(sub_anns$selec, sub_anns$element, sep = "-")
    
    # print(cs$family[1])
    # if (!file.exists(file.path(ann_spec_path, sub_anns$family[1], gsub(".wav", "--p1.jpeg",x))))
    full_spectrograms(X = sub_anns, flim = frq_range, sxrow = 4, rows = 10, ovlp = 50, collevels = seq(-100, 0, 5), parallel = 1, overwrite = FALSE, dest.path = file.path(ann_spec_path, sub_anns$family[1]), song = "song", fast.spec = TRUE, horizontal = TRUE, pb = FALSE)
    }
    )


```

# Add new recordings
```{r, eval = FALSE}

new_data_no_cuts <- readxl::read_excel("./data/raw/Marcelo Files - no Production cut.xlsx")

new_data_cuts <- readxl::read_excel("./data/raw/Marcelo Files - Production cuts.xlsx")

new_data_cuts <- new_data_cuts[!new_data_cuts$`Scientific Name` %in% rec_data$species,] 

nrow(new_data_cuts)

new_data_no_cuts <- new_data_no_cuts[!new_data_no_cuts$`Parent Species` %in% rec_data$species,] 

nrow(new_data_no_cuts)

head(new_data_no_cuts)

new_data_no_cuts$source <- "not publication cut"

new_data_cuts$source <- "publication cut"

clm <-read.csv("./data/raw/NEW_Clements-Checklist-v2022-October-2022.csv")
clm.sp <- clm[clm$category == "species", ]


clm.sp$genus <- sapply(clm.sp$scientific.name, function(x) strsplit(x, " ")[[1]][1])

new_data_cuts$genus <- sapply(new_data_cuts$`Scientific Name`, function(x) strsplit(x, " ")[[1]][1])


new_data_cuts$family <- sapply(seq_len(nrow(new_data_cuts)), function(x){ fam <- clm.sp$family[clm.sp$genus == new_data_cuts$genus[x]][1]

fam <- if (length(fam) == 0) NA else strsplit(fam, " \\(")[[1]][1]

return(fam)
})

new_data_no_cuts$genus <- sapply(new_data_no_cuts$`Parent Species`, function(x) strsplit(x, " ")[[1]][1])

new_data_no_cuts$family <- sapply(seq_len(nrow(new_data_no_cuts)), function(x){ fam <- clm.sp$family[clm.sp$genus == new_data_no_cuts$genus[x]][1]

fam <- if (length(fam) == 0) NA else strsplit(fam, " \\(")[[1]][1]

return(fam)
})


new_data_cuts$Orginal.Scientific.Name <- new_data_cuts$Scientific.Name <- new_data_cuts$`Scientific Name`

new_data_no_cuts$Orginal.Scientific.Name <- new_data_no_cuts$`Parent Species`

new_data_no_cuts$`Scientific Name` <- new_data_cuts$`Scientific Name` <- NULL

new_data_no_cuts$Scientific.Name <- new_data_no_cuts$`Parent Species`

new_data_no_cuts$file_url <- paste0("https://macaulaylibrary.org/asset/", new_data_no_cuts$`ML Catalog Number`)

new_data_cuts$file_url <- paste0("https://macaulaylibrary.org/asset/", new_data_cuts$`ML Catalog Number`)

new_data_no_cuts$species_ebird_url <- paste0("https://ebird.org/species/", new_data_no_cuts$`eBird Species Code`)


new_data_cuts$species_ebird_url <- paste0("https://ebird.org/species/", new_data_cuts$SpeciesCode)
new_data_cuts$`Common Name` <- new_data_cuts$`English Name`

new_data_cuts$Behaviors <- NA

names(new_data_no_cuts)
names(new_data_cuts)

new_data_cuts$Recordist <- paste(new_data_cuts$FirstName, new_data_cuts$LastName)
new_data_cuts$Year <- new_data_cuts$Month <- new_data_cuts$Day <- new_data_cuts$Locality <- new_data_cuts$`Loc ID`<- new_data_cuts$`Country State County` <- new_data_cuts$`Taxon Category` <- NA
    
common_colums <- intersect(names(new_data_no_cuts), names(new_data_cuts))

second_batch <- rbind(new_data_no_cuts[, common_colums], new_data_cuts[, common_colums])

second_batch$sound.files <- paste0(gsub(" ", "_", second_batch$Scientific.Name), "-ML", second_batch$`ML Catalog Number`, ".wav")


# ohun::feature_acoustic_data(path = "~/Downloads/combined/")
# warbleR::info_sound_files(path = "~/Downloads/combined/")


# this code is not organized (is a mess!)
fls <- list.files(path = "~/Downloads/combined/")

table(substr(fls, nchar(fls) - 3, nchar(fls)))

second_batch <- second_batch[second_batch$`ML Catalog Number` %in% gsub(".m4a|.mp3|.wav", "", fls),]

nrow(second_batch)

sub_fls <- fls[gsub(".m4a|.mp3|.wav", "", fls) %in% second_batch$`ML Catalog Number`]


table(substr(sub_fls, nchar(sub_fls) - 3, nchar(sub_fls)))


sum(!second_batch$`ML Catalog Number` %in% gsub(".m4a|.mp3|.wav", "", fls))

mp3_2_wav(samp.rate = 44.1, bit.depth = 16, path = "~/Downloads/combined/", overwrite = TRUE, dest.path = "~/Downloads/combined/")

# delete mp3s


setwd("~/Downloads/combined/")
for(i in grep("m4a$", fls, value = TRUE))
{
    cll <- paste0("ffmpeg -i ", i, " ", gsub("m4a$", "wav", i))
    system(cll)
}  

write.csv(second_batch, "./data/raw/sound_files_and_extended_metadata_second_batch.csv", row.names = FALSE)

second_batch <- second_batch[, c("family", "Scientific.Name", "species_ebird_url", "ML Catalog Number", "sound.files", "Common Name", "Orginal.Scientific.Name", "Behaviors", "file_url")]

write.csv(second_batch, "./data/raw/sound_files_metadata_second_batch.csv", row.names = FALSE)

fix_wavs(samp.rate = 44.1, bit.depth = 16, path = "~/Downloads/combined/")


fr <- file.rename(
    from = file.path("~/Downloads/combined/converted_sound_files", paste0(second_batch$`ML Catalog Number`, ".wav")), 
    to = file.path("~/Downloads/combined/converted_sound_files", second_batch$sound.files))

nrow(second_batch)
all(fr)


for(i in na.omit(unique(second_batch$family))){
  print(i)  
    if (!dir.exists(file.path("/run/user/1000/gvfs/smb-share:server=cinnas.local,share=neurobiología/marcelo_araya/bird_song_evolution/by_family", i)))
    dir.create(file.path("/run/user/1000/gvfs/smb-share:server=cinnas.local,share=neurobiología/marcelo_araya/bird_song_evolution/by_family", i))
    
    frm <- file.path("~/Downloads/combined/converted_sound_files", na.omit(second_batch$sound.files[second_batch$family == i]))
    tu <- file.path("/run/user/1000/gvfs/smb-share:server=cinnas.local,share=neurobiología/marcelo_araya/bird_song_evolution/by_family", i, na.omit(second_batch$sound.files[second_batch$family == i]))
    
    fr <- file.copy(frm, to = tu)
}    

```


<!-- light green box -->

<div class="alert alert-success">

# Takeaways {.unnumbered .unlisted}

- Doing good progress

</div>

&nbsp;

<!-- '---' adds a gray vertical line -->

---

&nbsp; 
 
 <!-- add packages used, system details and versions  -->
 
# Session information {.unnumbered .unlisted}

```{r session info, echo=F}

sessionInfo()

```
